---
title: "My first R Markdown report"
author: "Add your name here"
output: html_document
editor_options: 
  chunk_output_type: console
---

# R markdown file

This an R Markdown file. It combines text with code. This is text written in plain markdown and you can use markdown syntax to highlight text in **bold**, *italic* or _underlined_.

# Tasks

I have listed tasks for you below. I am also going to show you how to solve the task on the shared screen. You can follow along as I speak. 

## Your turn 1 - load R packages

1. Let's add a code chunk by clicking on the green C+ at the top of this window
2. Inside the code chunk, write: library(tidyverse)
3. Click on the green play button in the top right of the code chunk you have created

```{r}

library(tidyverse)

```

## Your turn 2 - Load JMP WASH data

1. Add a new code chunk by clicking on the green C+ at the top of this window
2. Inside the code chunk, write: read_csv("data/washdata.csv")
3. 

```{r}

read_csv("exercise/washdata.csv")

```

# Working with data

```{r}

washdata %>% 
  filter(iso3 == "UGA") %>% 
  filter(year == 2017) %>% 
  filter(residence == "national") %>% 
  filter(service == "sanitation")

```

