---
title: "Report on Water, Sanitation and Hygiene"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Your turn

The code for this exercise is already written. A lot of what is written in here will look foreign to you, but if you knit the document and inspect the code together witht plots that are created, then you may be able to identify how some parts of the code relate to the output.

1. Knit the document by clicking on the Knit Button on the top of this window
2. Inspect the document in the window on the bottom right


# Load R packages and import data

```{r}

# load libraries
library(tidyverse)
library(wbstats)

# import data
washdata <- read_csv("data/washdata.csv")

```

# Report

You work in Uganda and every 5 years your prepare a report for the East African Summit on WASH on how Uganda is performing on WASH.

In previous years, you have downloaded the helpful plots from the washdata.org website and compiled them into a short report.

You realize that you are performing the same task every year. Maybe there is another way to do this more efficiently.

```{r}


```

```{r}

# Pre-select your country, year and residence and service here

washdata_uga <- washdata %>%
  filter(iso3 == "UGA") %>% 
  filter(year == 2017) %>% 
  filter(residence == "national") 
```

# Global analysis

# Country analysis

## Water

```{r}

washdata_uga %>% 
  filter(service == "water") %>% 
  filter(indicator_type == "water_ladder") %>% 
  
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  theme_minimal()

```

## Sanitation

```{r}

washdata_uga  %>% 
  filter(service == "sanitation") %>% 
  filter(indicator_type == "sanitation_ladder") %>% 
  
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  theme_minimal()

```

## Hygiene

```{r}

washdata_uga %>% 
  filter(service == "hygiene") %>% 
  filter(indicator_type == "hygiene_ladder") %>% 
  
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  theme_minimal()

```

```{r child = "setup/setup.Rmd"}

```
