---
title: "Report on WASH Ladders"
author: "Add your name here"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
    theme: flatly
  pdf_document:
    toc: yes
subtitle: For `r params$name` on a `r params$residence` level in `r params$year`
params:
  year: 2017
  name: Uganda
  residence: national
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}

# I defined some settings about the appearance of the document here
# The things that are defined here are applied the whole document

knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE, 
                      eval = TRUE,
                      fig.width = 9, 
                      fig.asp = 0.5,
                      comment = "", 
                      cache = FALSE, 
                      out.width = "100%",
                      fig.align = "center",
                      dpi = 300,
                      message = FALSE)

```


# Your turn

The code for this exercise is already written. A lot of what is written in here will look foreign to you, but if you knit the document and inspect the code together with the plots and tables that are created, then you may be able to identify how some parts of the code relate to the output.

1. Knit the document by clicking on the Knit Button on the top of this window
2. Inspect the document that opens
3. Go back to the top of the document and change the name to a country name of your choice
4. Knit the document by clicking on the Knit Button on the top of this window 

# Parameterised reports

Parameterised reports allow you to specify one or more parameters to customise your analysis. I have pre-defined some parameters at the top of this document.

Learn more: https://bookdown.org/yihui/rmarkdown/parameterized-reports.html

# Load R packages and import data

We are loading the required R packages and import data.

```{r}

# load libraries
library(tidyverse)

# import data
washdata <- read_csv("data/washdata.csv")

```

The ISO3 country code, the year and residence are pre-selected here using parameters from the top of the document.

```{r}

# smaller subset of data is used. Here is where we use parameters defined at the top of the document
washdata_sml <- washdata %>%
  filter(name == params$name) %>% 
  filter(year == params$year) %>% 
  filter(residence == params$residence) 
  
# identify country name for selected ISO3 code
country_name <- washdata_sml %>% 
  select(name) %>% 
  distinct(name) %>% 
  pull(name)

```

# Data transformation

Here, I import the JMP color palette for the different service levels. I also prepare two objects that will set the order for service level variables. Both colors and factors levels are defined here, stored as an object in the Environment and are then re-used later.

```{r}

# Colors
color_scale_hygiene <- c("#ffd02e", "#fff381", "#b450c3")
color_scale_water <- c("#59caf8", "#fff381", "#ffda5a", "#ffbc02")
color_scale_sanitation <- c("#8cce8f", "#fff381", "#ffda5a", "#ffbc02")

# Factor level
fct_water <- c("At lest basic", "Limited (more than 30 mins)", "Unimproved", "Surface water")
fct_sanitation <- c("At least basic", "Limited (shared)", "Unimproved", "Open defecation")
fct_hygiene <- c("Basic", "Limited (without water or soap)", "No facility")

```


# WASH Report for `r country_name`

## Drinking Water Ladder

This plot shows the JMP ladder for access to drinking water. Safely managed drinking water is not displayed here.  

```{r}
# data transformation
washdata_sml %>% 
  filter(service == "water") %>% 
  filter(indicator_type == "water_ladder") %>%  # this sign connects the two
  
# data visualisation
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = color_scale_water) +
  labs(
    title = "Hygiene",
    subtitle = "Service levels do not display safely managed",
    caption = "Data source: washdata.org",
    x = NULL, 
    y = "Percentage",
    fill = "Service levels"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) 

```

## Sanitation Ladder

This plot shows the JMP ladder for access to sanitation. Safely managed sanitation is not displayed here.  

```{r}

# data transformation
washdata_sml  %>% 
  filter(service == "sanitation") %>% 
  filter(indicator_type == "sanitation_ladder") %>% 
  mutate(indicator = factor(indicator, levels = fct_sanitation)) %>%  # this sign connects the two
  
# data visualisation
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = color_scale_sanitation) +
  labs(
    title = "Saniation",
    subtitle = "Service levels do not display safely managed",
    caption = "Data source: washdata.org",
    x = NULL, 
    y = "Percentage",
    fill = "Service levels"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) 

```

## Hygiene Ladder

This plot shows the JMP ladder for access to hygiene.

```{r}

# data transformation
washdata_sml %>% 
  filter(service == "hygiene") %>% 
  filter(indicator_type == "hygiene_ladder") %>% 
  mutate(indicator = factor(indicator, levels = fct_hygiene)) %>%   # this sign connects the two
  
# data visualisation
  ggplot(aes(x = service, y = percent, fill = indicator)) +
  geom_col() +
  scale_fill_manual(values = color_scale_hygiene) +
  labs(
    title = "Hygiene",
    caption = "Data source: washdata.org",
    x = NULL, 
    y = "Percentage",
    fill = "Service levels"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(), 
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) 

```

